*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-23 Code/Manuscript_O2.Rmd
Added Naaman's bioinformatic methods

Near the equator, photoperiod remains relatively constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the year. The effective length of the photoperiod does, however attentuate with depth as dawn and dusk light at depth drops below the level needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions  [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes both at surface and at depth, with potentially complex effects upon viable growth niches [@liDiatomGrowthResponses2017; @prezelinDielPeriodicityPhytoplankton1992]. For example, Vaulot et al. @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.  To our knowledge, no study has as yet addressed *P. marinus* growth responses in relation to a  range of photoperiods.

@@

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using the Tidyverse package [@R-tidyverse] in RStudio v1.2.5019 (https://rstudio.com/)  (LOCATION OF CODE ON GITHUBO. The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b6f (Cytb6f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Terminal Oxidase (PTOX); plastocyanin (PC); ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen Prochlorococcus strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values file  (CITATION of FILES LOCATION).

@@

NAAMAN

@@

## **Prochlorococcus* genomic capacities for electron dissipation and DNA repair

>>>>

Near the equator, photoperiod remains relatively constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the year. The effective length of the photoperiod does, however attentuate with depth as dawn and dusk light at depth drops below the level needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions  [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes both at surface and at depth, with potentially complex effects upon viable growth niches [@liDiatomGrowthResponses2017; @prezelinDielPeriodicityPhytoplankton1992]. For example, Vaulot *et al*. (1995) @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.  To our knowledge, no study has as yet addressed *P. marinus* growth responses in relation to a  range of photoperiods.

@@

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using the Tidyverse package [@R-tidyverse] in RStudio v1.2.5019 (https://rstudio.com/)  (LOCATION OF CODE ON GITHUB). The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b6f (Cytb6f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Terminal Oxidase (PTOX); plastocyanin (PC); ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen Prochlorococcus strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values file  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData).

@@

Using the dataset Omar *et al*. (2023) [@omarAnnotationGenesEncoding2023], we filtered the dataset for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in Prochlorococcus marinus (MED4, MIT9313, SS120, and NATL2A) in order to generate a subset containing only those orthologs encoding enzymes directly mediating O~2~ metabolism. We grouped orthologs together by EC number and their KO number (KO number) and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from Prochlorococcus were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. XXXX

@@

## **Prochlorococcus* genomic capacities for electron dissipation and DNA repair {.unnumbered}
